---
title: "Exploring General Social Survey Data in R"
author: "Michael Chirico"
date: "Compiled `r format(Sys.time(), '%B %d, %Y at %R')`"
output: 
  rmarkdown::html_document
---
## Task 1 : Preliminaries

```{r loading}
library(data.table)
library(haven)

rm(list = ls(all = TRUE))
gc()

#setDT() converts the result to a data.table --
#  since read_dta is not part of the data.table
#  package, it returns the data as a data.frame,
#  which (in my opinion) is more of a pain to deal with
gss = setDT(read_dta("~/Dropbox/3/gss7210_r2_small.dta"))
```

`read_dta` keeps the variable label from Stata under a thing called an attribute of each variable. So, for example, the variable `racdin`, which has a sort of cryptic name, is actually much easier to understand by examining:

```{r labels}
gss[ , attr(racdin, "label")]

#to view all of the atribute labels in the data set, do:
sapply(gss, attr, "label")
```

This is still a lot of information; to narrow things down, we'll use a regular expression, which you can think of as a more powerful version of Googling. We're looking for something about segregation -- so presumably the variable label for any relevant variable will at the very least have the characters `"seg"` somewhere in it:

```{r search_labels}
grep("seg", sapply(gss, attr, "label"), value = TRUE)

#similarly, the one relevant to inter-racial marriage should 
#  at a minimum contain the string "marr"
grep("marr", sapply(gss, attr, "label"), value = TRUE)

#Now that we've identified the relevant variables, we can explore them
gss[ , class(racseg)]
```

The `labelled` class is specific to the `haven` package.  Look up the `?labelled` help file to learn more. The  `"labels"` attribute tells us the mapping between the integer values the variable takes and what that integer stands for.

```{r seg_mar}
gss[ , attr(racseg, "labels")]
gss[ , attr(racmar, "labels")]

## Once we know what the values stand for, the one-way table
##   will make sense once we see it:
gss[ , table(racseg)]
gss[ , table(racmar)]
```

Note that there are >50,000 observations in the whole  data set, but not nearly that many in these tables. That's because these questions are plagued by  extensive non-response; also, they simply weren't asked in some years of the survey. To see the missing responses, use the `useNA` argument to `table`:

```{r missing_table}
gss[ , table(racseg, useNA = "ifany")]
gss[ , table(racmar, useNA = "ifany")]
```
  
## Task 2 : Segregation

Given what we learned above, we'll want to exclude non-respondents from our calculations; also, those who are anti-segregation correspond to those responding either `3` or `4` on `racseg`:

```{r segregation, results = "hide"}
gss[!is.na(racseg), anti_segregation := 
      racseg %in% c(3, 4)]
```

## Task 3: Interracial Marriage

Defining support for interracial marriage is nearly identical:

```{r inter.marriage, results = "hide"}
gss[!is.na(racmar), pro_interracial_marriage := racmar == 2]
```

## Task 4: Averages by Year

Luckily the year of the observation is simply stored in the variable `year`. Remembering that there are many missing observations in each of our indicators, we'll have to use the `na.rm` argument to handle these. By default, R returns `NA` if we try and take the mean of any variable that has a missing observation; if we set `na.rm = TRUE`, R will exclude all of the missing rows and take the average of the remaining rows.

```{r averages.by.year}
#.(a = mean(b)) will return a column named "a" which
#  is equal to the mean of the column b.
gss[ , .(pct_anti_segregation = 
           mean(anti_segregation, na.rm = TRUE), 
         pct_pro_interracial_marriage = 
           mean(pro_interracial_marriage, na.rm = TRUE)),
     #recall that we use keyby to be sure the output is sorted.
     keyby = year]
```

## Task 5: Plot

```{r graph, results = "hide"}
#the first part is just a copy-paste of the
#  answer to the prior task
gss[ , .(pct_anti_segregation = 
           mean(anti_segregation, na.rm = TRUE), 
         pct_pro_interracial_marriage = 
           mean(pro_interracial_marriage, na.rm = TRUE)),
     #we take the result of the earlier task, and perform
     #  more calculations on the result by adding a 
     #  second pair of square braces; the second pair of
     #  square braces allows us to access the results of
     #  the first pair of square braces.
     keyby = year][ , {
       #see ?matplot
       matplot(year, cbind(pct_anti_segregation, 
                           pct_pro_interracial_marriage),
               main = "Evolution of Attitudes Towards Racial Issues",
               xlab = "Year", ylab = "Proportion", las = 1L,
               type = "l", lwd = 3L, lty = 1L, col = c("red", "blue"))
       legend("topleft", legend = c("Anti-Segregation",
                                    "Pro-Interracial Marriage"),
              col = c("red", "blue"), lty = 1L, lwd = 3L)}]
```
