---
title: "Plotting the Bernoulli Likelihood in R"
author: "Michael Chirico"
date: "Compiled `r format(Sys.time(), '%B %d, %Y at %R')`"
output: 
  rmarkdown::html_document
---

## Simulate Drawing 5 observations from Bernoulli(.45)

```{r sampling}
# for reproducibility
set.seed(20161021)

# Could also use rbinom(5, 1, .45)
x = runif(n = 5) <= .45

x
```

## Likelihood

The likelihood of a Bernoulli sample is given by:

$$ \mathcal{L}(p | x) = \prod_{i = 1}^n p^{x_i} (1 - p)^{1 - x_i} = p^{\sum_{i = 1}^n x_i} (1 - p)^{\sum_{i = 1}^n(1 - x_i)} $$

For example, suppose the first three observations were equal to `TRUE` and the last two were equal to `FALSE`. Each of the `TRUE` observations occurs with probability $p$; the others with probability $1-p$. The $\sum_{i = 1}^n x_i$ part counts the `TRUE` observations; the $\sum_{i = 1}^n (1 - x_i)$ part counts the `FALSE` observations.

We can code this up as:

```{r plotting}
#create a function or easier replication/repetition
bern.lik = function(p, x) p^sum(x) * (1-p)^sum(!x)

#create a grid of potential p values
ps = seq(0, 1, length.out = 1000L)

#store the likelihood values
plot(ps, bern.lik(ps, x), type = "l", lwd = 3L, las = 1L,
     xlab = "p", ylab = "Likelihood", col = "darkgreen",
     main = paste0("Likelihood Function on Size-5 Bernoulli Sample\n",
                   "With True Parameter .45"))
```

## Repeating for 10 observations

Just copy-paste, only changing what needs to be changed

```{r size_10}
n = 10
x = runif(n) <= .45
plot(ps, bern.lik(ps, x), type = "l", lwd = 3L, las = 1L,
     xlab = "p", ylab = "Likelihood", col = "darkgreen",
     main = paste0("Likelihood Function on Size-", n, " Bernoulli Sample\n",
                   "With True Parameter .45"))
```

## For 100 observations

```{r size_100}
n = 100
plot(ps, bern.lik(ps, runif(n) <= .45), type = "l", lwd = 3L, las = 1L,
     xlab = "p", ylab = "Likelihood", col = "darkgreen",
     main = paste0("Likelihood Function on Size-", n, " Bernoulli Sample\n",
                   "With True Parameter .45"))
```

## For 1000 observations

```{r size_1000}
n = 1000
plot(ps, bern.lik(ps, runif(n) <= .45), type = "l", lwd = 3L, las = 1L,
     xlab = "p", ylab = "Likelihood", col = "darkgreen",
     main = paste0("Likelihood Function on Size-", n, " Bernoulli Sample\n",
                   "With True Parameter .45"))
```

We could have done this all in a few lines of code like so:

```{r all_together}
par(mfrow = c(2L, 2L))
for (n in c(5L, 10L, 100L, 1000L)) {
  plot(ps, bern.lik(ps, runif(n) <= .45), type = "l", lwd = 3L, las = 1L,
     xlab = "p", ylab = "Likelihood", col = "darkgreen",
     main = paste0("Sample Size: ", n))
}
```

