---
title: "Simulating Election Results"
author: "Michael Chirico"
date: "Compiled `r format(Sys.time(), '%B %d, %Y at %R')`"
output: 
  rmarkdown::html_document
---

# The New York Times Model (Uncorrelated Polls)

Draws from the Beta distribution can be retrieved from `rbeta`. 

```{r nyt}
set.seed(20161104)
nn = 1e4 #10,000 draws
beta_draws = cbind(WI = rbeta(nn, shape1 = 102.922, shape2 = 85.1391),
                   MI = rbeta(nn, shape1 = 167.368, shape2 = 144.342),
                   PA = rbeta(nn, shape1 = 248.777, shape2 = 222.201))
#note the effect of having specified WI = ...
#  each column is clearly associated with a state
head(beta_draws)
```

## Generating shocks with a specified covariance

See [this](http://stats.stackexchange.com/questions/12953/generating-values-from-a-multivariate-gaussian-distribution) Q & A on CrossValidated.

Also scan the Wikipedia article about [Cholesky decomposition](https://en.wikipedia.org/wiki/Cholesky_decomposition) of a matrix.

The main takeaway is that if we want to generate a vector of random draws from the distribution $Y \sim N(\mu, \Sigma)$, we can transform an equal-length set of independent draws from $Z \sim N(0, I)$ with:

$Y = \mu + L Z$

Where $L$ is the Cholesky decomposition of $\Sigma$, i.e., $LL^{T} = \Sigma$. We can get the Cholesky decomposition of a matrix in R by using `chol`.

```{r nyt_disturbances}
#choosing the desired bias and variance-covariance matrix:
mu = c(WI = 0, MI = 0, PA = 0)
Sigma = matrix(c(.1, 0, 0,
                 0, .1, 0,
                 0, 0, .1),
               #if we don't specify nrow &/or ncol,
               #  R assumes we just want a vector
               nrow = 3L, ncol = 3L, 
               #row names come first, then column names;
               #  see ?matrix
               dimnames = list(c("WI", "MI", "PA"),
                               c("WI", "MI", "PA")))

nyt_err = mu + matrix(rnorm(3*nn), ncol = 3L) %*% chol(Sigma)
head(nyt_err)
#These numbers are too big! If we're adding numbers like 36%
#  or more we're bound to get outcomes outside the [0,1] range.
#  So we've got to try specifying Sigma again
Sigma = matrix(c(.01, 0, 0,
                 0, .01, 0,
                 0, 0, .01),
               #if we don't specify nrow &/or ncol,
               #  R assumes we just want a vector
               nrow = 3L, ncol = 3L, 
               #row names come first, then column names;
               #  see ?matrix
               dimnames = list(c("WI", "MI", "PA"),
                               c("WI", "MI", "PA")))

nyt_err = mu + matrix(rnorm(3*nn), ncol = 3L) %*% t(chol(Sigma))
```

## NYT Outcomes

```{r nyt_outcome}
nyt_out = beta_draws + nyt_err
head(nyt_out)
#Make sure that all of our outcomes
#  for each election make sense
apply(nyt_out, 2L, range)

hillary_wins_state = nyt_out > .5
head(hillary_wins_state)

hillary_states = rowSums(hillary_wins_state)
head(hillary_states)

hillary_win_pct = mean(hillary_states >= 2)
hillary_win_pct
```

# 538 Simulations

We can re-use `beta_draws` from above; just need to re-draw the error terms with a new `Sigma` matrix of our choosing:

```{r five38}
Sigma538 = matrix(c(.01, .00005, .00005,
                    .00005, .01, .00005,
                    .00005, .00005, .01),
                  nrow = 3L, ncol = 3L, 
                  dimnames = list(c("WI", "MI", "PA"),
                                  c("WI", "MI", "PA")))

five38_err = mu + matrix(rnorm(3*nn), ncol = 3L) %*% t(chol(Sigma538))
five38_out = beta_draws + five38_err
apply(five38_out, 2L, range)
hillary_win_pct = 
  mean(rowSums(five38_out > .5) >= 2)
hillary_win_pct
```

# Making this easier to tinker with

It's a pain to write out all of $\Sigma$, for two reasons.

 1. We know $\Sigma$ is symmetric, but we still have to type out all of the entries (wasted typing).
 2. It's a requirement that $\Sigma$ be positive semi-definite (a multivariate version of requiring that the variance is positive, see [here](https://en.wikipedia.org/wiki/Positive-definite_matrix)). This is a lot harder to eyeball; basically, things are driven by the relationship:
 
 $$ \rho_{XY} = \frac{\sigma_{XY}}{\sigma_X\sigma_Y} $$
 
 Since this can be at most one in absolute value, we know that an off-diagonal entry must be bounded between:
 
 $$ -\sigma_X \sigma_Y \leq \sigma_{XY} \leq \sigma_X \sigma_Y $$

It would be much easier if we could just specify the correlations, since it's easy to pick correlations, and the standard deviations, since it's easy to pick those, too.

So what we'll do is write a function that takes 3 things as input:

 1. The three means of the error terms
 2. The three standard deviations of the error terms
 3. The 3 correlations -- WI-MI, WI-PA, and MI-PA
 
 And outputs the NYT simulation result compared to the 538 simulation result.
 
 Here's the code; check it out on the widget here:
 
 
```{r as_function}
compare_simulations = function(mus, sds, corrs) {
  beta_draws_bi =
    cbind(rbeta(nn, shape1 = 102.922, shape2 = 85.1391),
          rbeta(nn, shape1 = 167.368, shape2 = 144.342),
          #include mean-shift here for minor efficiency boost
          rbeta(nn, shape1 = 248.777, shape2 = 222.201)) + mus
  
  #draw the same errors for both approaches, for efficiency
  ind_err = matrix(rnorm(3*nn), ncol = 3L)
  
  #NYT vs. 538 simulations differ only in
  #  off-diagonal terms of Sigma. So declare
  #  the Cholesky decomposition from the
  #  correlation matrix having 0 off-diagonal first
  corr_mat = diag(3)
  L = t(chol(diag(sds) %*% corr_mat %*% diag(sds)))
  nyt = 100*mean(rowSums((beta_draws_bi + (ind_err %*% L)) > .5) >= 2)
  # now add the specified correlations in the off-diagonal entries
  corr_mat[lower.tri(corr_mat)] = 
    corr_mat[upper.tri(corr_mat)] = corrs
  L = t(chol(diag(sds) %*% corr_mat %*% diag(sds)))
  five38 = 100*mean(rowSums((beta_draws_bi + (ind_err %*% L)) > .5) >= 2)
  cat("NYT says Hillary wins ", round(nyt), "% of the time.\n",
      "538 says Hillary wins ", round(five38), "% of the time.\n", sep = "")
}
```
