---
title: "PS2 Through R"
author: "Michael Chirico"
date: "Compiled `r format(Sys.time(), '%B %d, %Y at %R')`"
output: 
  rmarkdown::html_document
---

## Problem 1

We can just do this directly, or we can make a function to make our code more concise. Done directly:

```{r 1_nofunc}
#Part (d)
sum(exp(-4) * 4^(0:3)/factorial(0:3))

#Part (e)
sum(exp(-4) * 4^(11:12)/factorial(11:12))

#Part (f)
1 - sum(exp(-4) * 4^(0:5)/factorial(0:5))
```

If we wanted to do this, we might quickly tire of copy-pasting and/or writing out all of that code over and over. That's why functions were invented:

```{r 1_func}
#First, define a function
prob_days <- function(x) sum(exp(-4) * 4 ^ x / factorial(x))

#Now, the other parts become much more concise and readable:
prob_days(0:3)
prob_days(11:12)
1 - prob_days(0:5)
```

We could also have defined our function more generally (wherever 4 appears in the formula, it could be replaced by any positive number). Actually, this PMF is called the Poisson PMF, and since it's so common, R has two built-in function for it called `dpois` (giving the "density", i.e., the value of the PMF) and `ppois` giving the CDF. We could have used this like so:

```{r 1_pois}
ppois(3, lambda = 4)
sum(dpois(11:12, lambda = 4))
1 - ppois(5, lambda = 4)
```

## Problem 2

As above, we can either define a function or just do all the computations directly. First, without a special function:

```{r 2_nofunc}
#(c) (i)
sum(0:4 * choose(4, 0:4) * .55^(0:4) * (1-.55)^(4:0))

#(c) (ii)
choose(4, 0) * (.55)^0 * (1-.55)^4

#(c) (iii)
sum(choose(4, 3:4) * .55^(3:4) * (1-.55)^(4 - 3:4))

#(c) (iv)
20 * sum(choose(4, 0:2) * .55^(0:2) * (1-.55)^(4 - 0:2)) - 
  15 * sum(choose(4, 3:4) * .55^(3:4) * (1-.55)^(4 - 3:4))

#(c) (v)
100 * sum(choose(4, 3:4) * .55^(3:4) * (1-.55)^(4 - 3:4)) + 
  50 * choose(4, 2) * .55^(2) * (1-.55)^(4 - 2) - 
  200 * sum(choose(4, 0:1) * .55^(0:1) * (1-.55)^(4 - 0:1))
```

With a function:

```{r 2_func}
repub_prob <- function(x) choose(4, x) * .55 ^ x * (1 - .55) ^ (4 - x)

sum(0:4 * repub_prob(0:4))
repub_prob(0)
sum(repub_prob(3:4))
sum(c(20, 20, 20, -15, -15) * repub_prob(0:4))
sum(c(-200, -200, 50, 100, 100) * repub_prob(0:4))
```

And once again, this PMF is common enough to warrant R having a set of built-in functions for interacting with it: importantly, `dbinom` and `pbinom` for the density (PMF) and probability (CDF), respectively. We can also, since we're dealing with a discrete distribution, just calculate the entire distribution, store it in a vector, and extract the parts we need for each question:

```{r 2_binom}
repub.pmf = dbinom(0:4, size = 4, prob = .55)
sum(0:4 * repub.pmf)
repub.pmf[1L]
sum(repub.pmf[4:5])
sum(c(20, 20, 20, -15, -15) * repub.pmf)
sum(c(-200, -200, 50, 100, 100) * repub.pmf)
```

## Problem 4

All limits can typically be guessed by plotting. For which R is perfect!

```{r 4_plots}
# two-sided limit, so take points on both sides
x <- seq(-6, -2, length.out = 1000L)

plot(x, (x^2 + 5*x + 4)/(x^2 + 3*x - 4),
     type = "l", lwd = 3L, xlab = "x",
     ylab = expression(f(x)), main = "4 (a)")
abline(v = -4, lty = 2, lwd = 2, col = "red")

# one-sided limit, so only take points from one side
x <- seq(2, 4, length.out = 1000L)
plot(x, sqrt(16 - x^2),
     type = "l", lwd = 3L, xlab = "x", 
     ylab = expression(f(x)), main = "4 (b)")
abline(v = 4, lty = 2, lwd = 2, col = "red")

# two-sided limit, so take points on both sides
x <- seq(-3, 1, length.out = 1000L)

plot(x, (x - 2)/(x^2 + 4*x - 3),
     type = "l", lwd = 3L, xlab = "x", 
     ylab = expression(f(x)), main = "4 (c)")
abline(v = -1, lty = 2, lwd = 2, col = "red")

# two-sided limit, so take points on both sides
x <- seq(-6, -2, length.out = 1000L)

plot(x, (1/4 + 1/x)/(4 + x),
     type = "l", lwd = 3L, xlab = "x",
     ylab = expression(f(x)), main = "4 (d)")
abline(v = -4, lty = 2, lwd = 2, col = "red")
```

## Problem 5

R has the ability to differentiate simple functions for you as well (admittedly, it's pretty rudimentary), using the `D` function -- see `?D`.

```{r 5_deriv}
D(expression(4*x^3 + 2*x^2 + 5*x + 11), "x")
D(expression(sqrt(30)), "x")
D(expression(2^3 + (sin(x))^3), "x")
D(expression(log(9*t + 1)), "t")
D(expression(x^3 * cos(11*x)), "x")
D(expression(log(x^2 * exp(x))), "x")
D(expression((1/y^2 - 3/y^4) * (y + 5*y^3)), "y")
D(expression(3*t - 1/2 * t + 1), "t")
```

## Problem 6

R can integrate as well! See `?integrate`.

```{r 6_integral}
integrate(function(x) x^(3/7), 0, 1)
integrate(function(x) 3/x^4 + 2, 1, 2)
integrate(function(x) 2^x, 8, 9)
integrate(function(x) sqrt(x^5 + 2), 3, 3)
```
