---
title: "Simulation of PS6:2"
runtime: shiny
output: html_document
---

For the purposes of this simulation, suppose $\theta_0 = 6$. That is, the true value of $\theta_0$ is known to us, and it's 6. Everything else will be a matter of trying to use the data we've got to identify this.

## Part c)

We can estimate the CDF for a given value of $n$ by repeatedly (say, $m$ times) taking uniformly distributed samples, and calculating $\hat{\theta}_n$ on each of these $m$ samples. This is made much easier in R using the `replicate` function. 

```{r part.c.code, eval = FALSE}
#fix n, and take m samples
theta_ns = replicate(m, 
                     #theta_n is, for each sample, the max of the draws
                     max(
                       #on each sample, generate n U[0, theta_0] draws
                       runif(n, min = 0, max = theta_0)
                     ))

#To get the CDF of this, we can do
ecdf(theta_ns)(theta_ns)
```


```{r part.c.plot, echo = FALSE}
#set "true" theta_0:

th = 6

inputPanel(
  sliderInput("n.c", label = "Sample Size (n)",
              min = 1, max = 500, value = 2, step = 1),
  
  selectInput("m.c", label = "Simulation Size (m)",
              choices = c(100, 1000, 10000), selected = 1000)
)

renderPlot({
  theta_ns = sort(replicate(input$m.c, max(runif(input$n.c, 0, th))))
  plot(theta_ns, ecdf(theta_ns)(theta_ns), type = "l", lwd = 3L, 
       main = expression(paste("Estimated CDF of ", hat(theta)[n])),
       cex = 2, cex.lab = 1.5, xlim = c(0, th),
       ylab = expression(paste("Pr. At Most ", theta)), xlab = expression(theta))
})
```

## Part d)

Again, we can just find $m$ samples, and on each of these, for a given $\epsilon$, calculate this probability as an expectation.

```{r part.d.code, eval = FALSE}
mean(abs(replicate(m, max(runif(n, 0, theta_0))) - theta_0) < epsilon)
```

```{r part.d.plot, echo = FALSE}
inputPanel(
  sliderInput("n.d", label = "Sample Size (n)",
              min = 1, max = 500, value = 2, step = 1),
  
  selectInput("m.d", label = "Simulation Size (m)",
              choices = c(100, 1000, 10000), selected = 1000)
)

renderPlot({
  par(mar = c(5.1, 5.1, 4.1, 2.1))
  theta_ns = replicate(input$m.d, max(runif(input$n.d, 0, th)))
  epsil = seq(0, th, length.out = 100)
  plot(epsil, sapply(epsil, function(ee) mean(abs(theta_ns - th) < ee)), 
       main = expression(paste("How Far ", hat(theta)[n], " Gets from ", theta[0])),
       cex = 2, cex.lab = 1.5, xlim = c(0, th), type = "l", lwd = 3L, 
       ylab = expression(paste("Pr[|", hat(theta)[n], "-", theta[0], "|<", epsilon, "]")),
       xlab = expression(epsilon))
})
```

## Part g)

Instead of using `ecdf`, we can use `density` to get the PDF:

```{r part.g.code, eval = FALSE}
density(replicate(m, max(runif(n, min = 0, max = theta_0))))
```

Note that, for small $n$, since it is rare to get a $\hat{\theta}_n$ close to the truth, our plot does a poor job of estimating the true PDF for such $\theta$.

```{r part.g.plot, echo = FALSE}
inputPanel(
  sliderInput("n.g", label = "Sample Size (n)",
              min = 1, max = 500, value = 2, step = 1),
  
  selectInput("m.g", label = "Simulation Size (m)",
              choices = c(100, 1000, 10000), selected = 1000)
)

renderPlot({
  dd = density(replicate(input$m.g, max(runif(input$n.g, 0, th))))
  plot(dd$x, dd$y, type = "l", lwd = 3L, 
       main = expression(paste("Estimated PDF of ", hat(theta)[n])),
       cex = 2, cex.lab = 1.5, xlim = c(0, th),
       ylab = "Density", xlab = expression(theta))
})
```

## Part j)

```{r part.j.plot, echo = FALSE}
inputPanel(
  sliderInput("n.j", label = "Sample Size (n)",
              min = 1, max = 500, value = 2, step = 1),
  
  selectInput("m.j", label = "Simulation Size (m)",
              choices = c(100, 1000, 10000), selected = 1000)
)

renderPlot({
  dd = density(replicate(input$m.j, 2*mean(runif(input$n.j, 0, th))))
  plot(dd$x, dd$y, type = "l", lwd = 3L, 
       main = expression(paste("Estimated PDF of ", ring(theta)[n])),
       cex = 2, cex.lab = 1.5, xlim = c(0, 2*th),
       ylab = "Density", xlab = expression(theta))
})
```

